# Basics

## Prerequisites

We introduce the basics in R programming in this chapter. We will review the basic operators and data types in this chapter. We also provide an introduction to the basic R data structures (including tibbles and data tables).

Most of this chapter involves working with R basic operators and data types, which do not require any extra packages. We will also introduce the **tibble** package, which forms part of the tidyverse package in section \@ref(tibbles), and the  **data.table** package in section \@ref(data-table)

```{r, message=FALSE}
library(tidyverse)
library(data.table)
```

## Basic operators

Basic arithmetic operators (`+`, `-`, `*`, `/`, `^`, `%') would work like your calculator

```{r}
3 + 2 # addition
3 - 2 # subtraction
3 * 2 # multiplication
3 / 2 # division
3^2 # exponent
3 %% 2 # mod (remainder of a division)
```


R uses the `<-` operator for assignments. You can read the following code as assigning the outcome of `2 + 3`, `5`, and `3` to the object `value_a`,`value_b`, and `value_c` respectively, which stores it for later use. 

```{r}
value_a <- 2 + 3
value_b <- 5
value_c <- 3
```

You can `print` what is stored in the object to console with

```{r}
print(value_a)
```

You can use relational operators to compare how one object relates to another. 

```{r}
2 + 3 == 5 # TRUE that 2 + 3 equals 5
2 + 3 != 5 # FALSE that 2 + 3 not equals to 5
2 + 3 < 3 # FASLE that 2 + 3 is less than 3
2 + 3 > 3 # TRUE that 2 + 3 is more than 3
2 + 3 <= 5 # TRUE that 2 + 3 is less than or equal to 5
2 + 3 >= 5 # TRUE that 2 + 3 is more than or equal to 5
```

You can use logical operators to connect two or more expressions. For example, to connect the results of the comparisons made using relational operators. 

```{r}
(2 + 3 == 5) & (2 + 3 < 3) # AND operator
(2 + 3 == 5) | (2 + 3 >= 3) # OR operator
!(2 + 3 == 5) # NOT operator
```



## Basic data types 

There are basic data types (also known as atomic data types) in R in order to use them.

Data Type      Examples                   Additional Information         
-----------    ---------------------      -----------------------------------------------
Logical        `TRUE`, `FALSE`            Boolean values 
Numeric        `1`, `999.9`               Default data type for numbers
Integer        `1L`, `999L`               `L` is used to denote an integer
Character      `"a"`, `"R for BES"`       Data type for one or more characters
Complex        `2 + 3i`                   Data type for numbers with a real and imaginary component
Raw            `charToRaw("R for BES")`   Not commonly used data type used to store raw bytes


You can use the function `class()` to find out the class of any object in R.

```{r}
var_a <- 1
class(var_a)

var_b <- 1L
class(var_b)

var_c <- "R for BES"
class(var_c)
```


## Basic data structures

The basic data structures in R include factors, atomic vectors, lists, matrices, and `data.frame`s 

Factors are used in R to represent categorical variables. Although they appear similar to character vectors they are actually stored as integers. You can use the function `levels()` to output the categorical variables and `nlevels()` to check the number of categorical variables.

```{r create-factors}
eye_color <- factor(c("brown", "black", "green", "brown", "black", "blue"))
nlevels(eye_color)
levels(eye_color)
```

Atomic vectors or more frequently referred to as vectors are a data structure that is used to store multiple objects of the same data type (logical, numeric, integer, character, complex, or raw). Vectors are one-indexed (i.e., the first element is indexed using `[1]`) and you can get the number of elements in the vector using the function `length()`.

```{r create-vectors}
vec_num <- c(1, 2, 3, 4)
class(vec_num)

vec_char <- c("R", "for", "BES")
class(vec_char)

# coercion if data types are mixed
vec_mix <- c("R", 4, "BES")
class(vec_mix)

# you can easily combine vectors using the c function
c(vec_char, vec_mix)

# vector length
length(vec_num)

# access first element of vector
vec_num[1]
```


Lists are an ordered data structure that is used to store multiple R objects of different types. The function `list()` is used to create a list and a list in R can be accessed using a single `[]` or double brackets `[[]]`. Using `[]` returns a list of the selected element while using `[[]]` returns the selected element. Using the function `length()`, you can obtain the number of objects in a list.

```{r}
my_list <- list(
  c(1, 2, 3, 4),
  c("a", "b"),
  1L,
  matrix(1:9, ncol = 3)
)

my_list_a <- my_list[1]
class(my_list_a)

my_list_b <- my_list[[1]]
class(my_list_b)

length(my_list)
```

If you have named the elements in your list, you could also access them by specifying their names in the brackets or using the `$` operator. 

```{r}
named_list <- list(
  a = c(1, 2, 3, 4),
  b = c("a", "b"),
  c = 1L,
  d = matrix(1:9, ncol = 3)
)

class(named_list["a"])

class(named_list[["a"]])

class(named_list$a)
```


A matrix is a two dimensional data structure that is used to store multiple objects. You can use the function `matrix()` to create a matrix using the `ncol` and `nrow` argument to specify the number of columns and rows respectively, and the `byrow` argument to specify how the data in would be ordered. 

```{r}
matrix(1:12, ncol = 3, byrow = FALSE)

matrix(1:12, nrow = 3, byrow = FALSE)

matrix(1:12, ncol = 3, byrow = TRUE)
```

Aside from numeric data types, a matrix can also be used to store other data types as long as they are homogeneous. To store heterogeneous data types, you should use a data frame which is introduced next. 

```{r}
matrix(c("brown", "black", "green", "brown", "black", "blue"), ncol = 2)

matrix(c("TRUE", "TRUE", "FALSE", "FALSE"), ncol = 2)

matrix(c(1L, 2L, 3L, 4L), ncol = 2)
```

A `data.frame` is a two-dimensional data structures that are used to store heterogeneous data types in R. As a result of it's convenience, data frames are a commonly used data structure in R. You can use the function `data.frame()` to create a data frame. 

```{r}
df <- data.frame(
  x = c(1, 2, 3),
  y = c("red", "green", "blue"),
  z = c(TRUE, FALSE, TRUE)
)
```


## Tibbles {#tibbles}

Tibbles are basically a modified version of R's `data.frame`. You can create a tibble using the function `tibble()`. Alternatively, you can coerce a data frame into a tibble using `as_tibble()`


```{r}
tibble(
  x = c(1, 2, 3),
  y = c("red", "green", "blue"),
  z = c(TRUE, FALSE, TRUE)
)

as_tibble(iris)
```

A key difference lies in how tibbles are printed. Printing a tibble only results in the first ten rows being displayed with an explicit reporting of each column's data type.

```{r}
tb <- as_tibble(iris)
print(tb)

df <- as.data.frame(iris)
print(df)
```

Unlike a `data.frame`, tibbles provides clarity on the data structure that it returns. When indexing with tibbles, `[` always return another tibble while `[[` and `$` alway returns a vector. In contrast, single column data frames are often converted into atomic vectors in R unless `drop = FALSE` is specified.

```{r}
class(tb[, 1])

class(tb[[1]])

class(tb$Sepal.Length)

class(df[, 1])

class(df[, 1, drop = FALSE])
```


Additionally, tibbles do not do partial matching and raises a warning unless the variable specified is an exact match.

```{r}
tb$Sepal.Length

df$Sepal.Lengt
```

You can read more about tibbles by typing vignette("tibble") in your console. 


## data.table {#data-table}

Like tibbles, `data.table`s are an enhanced version of `data.frame`s. You can create a `data.table` using the function `data.table()`. You can also coerce existing R objects into a `data.table` with `setDT()` for `data.frame`s and lists, and `as.data.table()` for other data structures. Note that `as.data.table()` also works with `data.frame`s and lists. However `setDT()` is more memory efficient because it does not create a copy of the original data frame or list but instead returns a data table by reference. 

```{r}
dt <- data.table(
  x = c(1, 2, 3),
  y = c("red", "green", "blue"),
  z = c(TRUE, FALSE, TRUE)
)

class(
  setDT(
    data.frame(c(1, 2, 3))
  )
)
```

`data.table`s provide additional functionality through the way it is queried. The general form for working with a data table is `[i, j, by]`, which can be read as subset rows using `i`, operate on `j`, and grouped by `by`. 

Lets see how this work using the `iris` example dataset.

```{r}
dt <- as.data.table(iris)

print(dt)
```

You can filter the rows to only contain `Species == "virginica`.

```{r}
dt[Species == "virginica"]
```

You can select the columns using the `j` expression. Note the change in the type of data structure when `list()` or `.()` is used.

```{r}
dt[, Sepal.Length]

class(dt[, Sepal.Length])

class(dt[, list(Sepal.Length)])

class(dt[, .(Sepal.Length)])
```

You can select multiple columns with `list()` or `.()`.

```{r}

dt[, .(Sepal.Length, Species)]
```


Aside from selecting columns using `j`, you can carry out computations on `j` involving one or more columns and a subset of rows using `i`.

```{r}

dt[, mean(Sepal.Length)]

dt[, .(
  Sepal.Length.Mean = mean(Sepal.Length),
  Sepal.With.Mean = mean(Sepal.Width)
)]

dt[
  Species == "virginica" & Sepal.Length < 6,
  .(
    Sepal.Length.Mean = mean(Sepal.Length),
    Sepal.With.Mean = mean(Sepal.Width)
  )
]
```

You can then use the `by` expression in data tables to perform computations by groups. 

```{r}

dt[, .(
  Sepal.Length.Mean = mean(Sepal.Length),
  Sepal.With.Mean = mean(Sepal.Width)
),
by = Species
]
```

The `.N` variable that counts the number of instances is particularly useful when combined with `by`.

```{r}

dt[, .N, by = Species]
```

You can also apply it to multiple columns using the `list()` or `.()` notation. You can read the code below as calculating the mean of `Speal.Length` and the number of instances (given by `.N`) grouped by their `Species` and whether `Sepal.Length < 6`. 

```{r}
dt[, .(Sepal.Length.Mean = mean(Sepal.Length), .N),
  by = .(Species, Sepal.Length < 6)
]
```

`data.table`s add, update, and delete columns by reference to avoid redundant copies for performance improvements. You can use the `:=` operator to add, update, and delete columns in `j` by reference. There are two forms for using `:=` and they are: `[, LSH := RHS]` and `[, `:=` (LHS = RHS)]`. 

```{r}
dt <- as.data.table(iris)

dt[, Sepal.Sum := .(Sepal.Length + Sepal.Width)]
head(dt)

dt[, `:=`(Petal.Sum = Petal.Length + Petal.Width)]
head(dt)
```

Note that in the above code, we do not need to make any assignments back to a variable because the modification is done by reference or in place. In other words we are modifying `dt` and not a copy of `dt`. Therefore, you will also see that if you run the entire code chunk above, `dt` will contain both columns `Sepal.Sum` and `Petal.Sum`.

Since `:=` is used in `j`, it can be combined with `i` and `by` as we have seen in the earlier parts of this sub-section. 

```{r}

dt[Species == "versicolor" | Species == "virginica", Sepal.Length := 0]
head(dt)

dt[, Sepal.Length.Mean := mean(Sepal.Length),
  by = .(Species)
]
head(dt)
```


You can find out more about `data.table`s by typing `vignette(package = "data.table")` into the console. 


