
# Dates and Times {#datetime}

## Prerequisites

In this chapter, we will introduce the **{lubridate}** package that is designed to make it easier to work with dates and times in R. 

```{r, message=FALSE}
library(lubridate)
```


## Parsing dates and times

You can easily transform dates and times stored as character vectors to `Date` or `POSIXct` (or date-time) objects in R using {lubridate}. All you need to do is specify the order of the year `y`, month `m`, and the day `d` and {lubridate} will automatically figure out the format.

```{r}
dmy("3/12/1985")
ymd("85/12/03")
mdy("Dec 3rd 1985")
ydm("85-3rd-december")
```

The same functions can also convert numeric vectors.

```{r}
ymd(19851203)
dmy(31285)
```

You can create Date objects with the time component using a `POSIXct` or date-time object simply by adding an underscore followed by the order of the hour `h`, minute `m` and second `s`. 

```{r}
ymd_h("1985/12/03 21")
ymd_hm("1985/12/03 21:05")
ymd_hms("1985/12/03 21:05:30")
```

You can also specify the time zone by providing inputs to the argument `tz`. You can find out more about time zones in R by typing `?timezones` into the console. 

```{r}
ymd_hms("1985/12/03 21:05:30", tz = "Singapore")
```

Date vs date-time objects that are created using {lubridate}. By default, dates will be created as `Date` objects without the time component. You can force the creation of a date-time object by including the timezone `tz` argument. 

```{r}
class(ymd("1985/12/03"))
class(ymd("1985/12/03", tz = "Singapore"))
class(ymd_hms("1985/12/03 21:05:30"))
```


## Extracting components

{lubridate} provides simple functions that allows you to easily get different components of a date or date-time object. These functions are especially useful when analyzing time-series data and when you want to group your data by a particular time period. 

```{r}

my_dt <- ymd_hms("1985/12/03 21:05:30")
year(my_dt)
month(my_dt)
day(my_dt)
hour(my_dt)
minute(my_dt)
second(my_dt)
wday(my_dt, label = TRUE)
```



