# Importing Data

```{r, include=FALSE}
source("_common.R", local = knitr::knit_global())
```

## Prerequisites

In this chapter, we focus on the use of the **{readr}** package that forms a
part of the **{tidyverse}** package. **{readr}** provides a fast and easy way to
parse a file because it uses a sophisticated parser that guesses the data type
for each column along with the flexibility to specify what to parse. You will
also need the **{here}** package (to easily provide relative file paths) and the
**{lubridate}** package (to work with dates and times).

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(here)
```

In this chapter we will also be using the following datasets. If the following
code returns `FALSE`, you should refer to the section [Finding your file] in the
[Preface] on where you can download the datasets.

```{r, eval=FALSE}
file.exists(here("data", "iris.csv"))
file.exists(here("data", "building.csv"))
```

## Finding your file

Before importing data, you need to first let R know where to find the file by
providing the file path. You can provide file paths relative to the top-level
directory of the current R project with the `here()` function.

```{r}
here()
```

To reference a file named `iris.csv` located in the `data` folder that is
located in the top-level directory of your project:

```{r}
here("data", "iris.csv")
```

## Parsing a csv file

We will focus on the `read_csv()` function because building data are typically
stored in `csv` format. You can find out more about other file formats in
[readr's documentation](https://readr.tidyverse.org).

The easiest way to parse a file is by providing the file path. In most cases, it
will just work as expected where `readr` correctly guesses the column
specification (that gets printed to the console) and you get a tibble as
specified.
[`readr_example()`](https://readr.tidyverse.org/reference/readr_example.html) is
a function that makes it easy to access example files to demonstrate `readr`'s
capabilities.

```{r}
read_csv(here("data", "iris.csv"))
```

## Building data

Data from buildings are often messy without a standardized format. Therefore, it
is not uncommon if `readr` does not guess correctly. Parsing date/times is
particularly important when working with time-series building data.

For example, with the `building.csv` dataset, you can see that there are
problems parsing the `timestamp` column. Specifically, column `timestamp` was
parsed as a character vector when it actually contains a date/time. The `skip`
argument is particularly useful since building data often contains meta-data
that you want to exclude. A character vector can also be supplied to specify the
column names.

```{r}

read_csv(
    here("data", "building.csv"),
    col_names = c(
        "datetime",
        "power"
    ),
    skip = 2
)
```

You can correct this by taking advantage of the {lubridate} package that was
covered in Chapter \@ref(datetime). To recap, you can transform a character
vector into a datetime object by specifying the order of the year `y`, month
`m`, day `d`, hour `h`, minute `m` and second `s` with the date (`y`, `m`, and
`d`) and the time (`h`, `m`, and `s`) separated by an underscore `_`.

```{r}
bldg <- read_csv(
    here("data", "building.csv"),
    col_names = c(
        "datetime",
        "power"
    ),
    skip = 2
)

# transform the datetime character vector into a datetime object using {lubridate}
bldg$datetime <- dmy_hm(bldg$datetime)
```
